Combination of field and experimental data together with computational models

### 2022-02-03 manuscript_1.0/manuscript_word.Rmd
Edit's related to Redouan's comments

---
title: Combination of field and experimental data together with computational models
  reveal cognitive mechanisms behind cleaner fish behaviour
author:
- Andrés E. Quiñones^[Institute of Biology, University of Neuchâtel, Neuchâtel, Switzerland
  - andresqp@gmail.com]
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  word_document: default
  pdf_document: default
fig_caption: yes
fontsize: 12pt
header-includes:
- \usepackage{float}
- \usepackage{setspace}\doublespacing
- \linespread{2}
- \usepackage{lineno}
- \linenumbers
- \newcommand{\beginsupplement}{ \setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}}
  \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
keep_tex: yes
linespread: 2
bibliography: Cleanerlearning.bib
style: M:\Zotero\styles\behavioral-ecology-and-sociobiology.csl
spacing: double
abstract: "While it is generally straightforward to quantify individual performance
  in cognitive experiments, identifying the underlying cognitive processes remains
  a major challenge. Often, different mechanistic underpinnings may yield similar
  performances, and Lloyd Morgan’s cannon warrants acceptance of the simpler explanation.
  Alternatively, when the different mechanisms interact with environmental conditions,
  variation in performance across environments might allow to statistically infer
  the mechanism responsible. However, to do this it is necessary to have quantitative
  predictions for the different candidate mechanisms. Here, we use a set of computational
  models to get quantitative predictions from alternative learning mechanisms, and
  using Bayesian statistics fit the model parameters to performance data. We used
  experimental data on performance in an ephemeral reward task by wild-caught cleaner
  fish Labroides dimidiatus, as well as cleaner and client fish densities from the
  locations of capture. The task can in principle be solved by estimating future consequences
  of an action, or by perceiving the removal of the ephemeral reward as psychological
  punishment (negative reinforcement). We found that a model where cleaners only estimate
  the future consequences of their actions explains best which cleaner relative abundances
  cause the fish to develop a preference for an ephemeral food source. This model
  also yields performances that can be considered the result of locally optimal decision-rules,
  in contrast to the negative reinforcement model.  We argue that the combination
  of computational models with data is a powerful tool to infer the mechanistic underpinning
  behind cognitive performance. \n"
keywords: learning, behaviour, cleaners, bayesian statisitics
---

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning = FALSE,message=FALSE)
library(here)
```

Cooperation among individuals of the same or different species is
ubiquitous in nature. Whenever such interactions involve some degree 
of partner choice, biological market theory predicts that the exchange 
rate of goods and/or services traded adjusts to the law of supply and demand 
[@noe_Biological_1995a]. While biological market models have largely focused 
on evolutionary time scales [@noe_Biological_1994a;@johnstone_Mutualism_2008a] 
$\color{red}{\text{Hoekstra xxxx}}$, supply and demand conditions, which typically 
depend on the abundance of the species involved, certainly vary in time 
and space within ecological time scales. Therefore, natural selection should 
favour the ability to flexibly adjust decisions and behavioural output to 
current market conditions. Indeed, such adjustments have been repeatedly 
documented [@axen_signalling_1996]. In animals, an obvious general candidate 
mechanism for the strategic adjustment is the cognitive machinery. However, 
it is not clear which cognitive mechanisms allow individuals to adjust their 
behaviour to the varying market conditions. As adjustments to changes in 
market conditions have been documented in a great variety of species with 
highly variable brain anatomies, the question arises to what extent mechanisms 
beyond basic associative learning may be involved.

One example of strategic adjustment in a biological market is the marine
cleaning mutualism involving the cleaner fish *Labroides dimidiatus* and
'client' fish. Clients offer themselves as food patches for cleaners to
remove ectoparasites, which gives the cleaners calories and the clients
improved health [@waldie_LongTerm_2011;
@ros_Does_2011;@triki_Effects_2016]. As cleaners have about 2000
interactions per day [@grutter_Relationship_1995], they regularly face
the situation that two clients seek cleaning simultaneously. In such
conditions, it is important to distinguish between 'resident' and
'visitor' client species [@bshary_Cleaner_2001a]. The residents have
small home ranges and hence access to only the local cleaner, while the
visitors have large home ranges and hence access to two or more
cleaners. Thus, visitors have the option to switch to another cleaner if
being made to wait, while residents must wait for inspection. Indeed,
visitors have been observed to use their partner choice options in that
way [@bshary_Choosy_2002], which may explain why cleaners give
visitors service priority in a field study in the Red Sea
[@bshary_Cleaner_2001a]. More generally, cleaners apparently prefer to
interact with visitors [@triki_Marine_2021]. Such a preference may be
linked to a client's value as a food patch; visitors are on average
larger than residents [ @bshary_Cleaner_2001a], and size is positively
correlated with parasite load [@grutter_Relationship_1995a]. However, 
early laboratory experiments showed that leaving causes service 
priority independent of client value [@bshary_Asymmetric_2002]. 
The experiments consisted of asimultaneous two-choice task, 
where cleaners had to decide which of two
Plexiglas plates, which offered equal amounts of food, to approach first
and eat mashed prawn of it. One plate was made to behave like a
resident, i.e. it would remain until the cleaner had eaten off it. The
other plate was made to behave like a visitor, i.e. it would be removed
if not inspected first. Cleaners learned to prefer the visitor plate and
hence obtained the double amount of food [@bshary_Asymmetric_2002].
Adult cleaner fish outperform various primates as well as rats and
pigeons in this original version of the biological market task
[@salwiczek_Adult_2012; @zentall_Early_2017]; only African grey
parrots solve this task as well [@pepperberg_Can_2014]. As the correct 
solution must be learned, and primates are known to readily distinguish 
between one and two reward units if presented in more conventional ways 
$\color{red}{\text{ref}}$, what makes the task difficult for most non-cleaners?

Recent research on the cognitive tool kit needed to solve an ephemeral 
reward task has used a broad approach as proposed by @shettleworth_Cognition_2009, 
who defined cognition as includes all ways in which animals take in information 
through the senses, process, retain and decide to act on it. Indeed, perception 
of relevant cues is of major importance for individual performance; 
i.e. what relevant information is encoded in the design of the two 
plates (colour versus size; [@wismer_Cuebased_2019]), or whether 
the information is in the colour of
the food [@pretot_Comparative_2021; @pretot_Comparing_2016], and
whether the food is visible or hidden/released only when the subjects
made choices on a touch screen [@pretot_Factors_2016]. However, identifying 
cues as salient is not the only challenge of the ephemeral reward task, 
as revealed by proximate learning models. Such models allow varying the 
cognitive tool kit and to evaluate which minimal kit is necessary to solve 
the task at hand (e.g. [@dubois_Model_2021]). Applied to the ephemeral 
reward task, learning models showed that basic reinforcement learning does 
not suffice to solve the task [@prat_Modelling_2022; @quinones_Reinforcement_2019]. 
This is particularly so when models assume the more complex natural 
situation in which cleaners not only face resident-visitor pairs but also 
visitor-visitor and resident-resident pairs, as well only a resident or 
only a visitor. To be able to give visitors priority over residents, 
cleaners need to be able to assess a client’s value separately for the 
three possible forms of encounter (alone, paired with a fish with the 
same strategic option, paired with a fish with the alternative strategic option) 
[@quinones_Reinforcement_2019]. The ability to keep such separate scores 
for the same stimulus has been termed configurational learning, 
or chunking, or segmentation (see references in @prat_Modelling_2022).

To solve the ephemeral reward task, cleaners also need to account for 
the future consequences of current decisions. In the model by 
Quiñones et al. @quinones_Reinforcement_2019, this could be achieved in 
two non-mutually exclusive ways: through low temporal discounting of 
future effects, also termed ‘chaining’ [@enquist_Power_2016]; and/or 
through perceiving a visitor client leaving as psychological punishment 
(i.e. as a negative reinforcer). Given that a negative reinforcer is part 
of the ubiquitous learning mechanism termed operant conditioning 
[@thorndike_Animal_1898; @skinner_Behavior_1938], this explanation 
could be considered the simpler one. Furthermore, accounting for the 
future effects of an action is considered an advanced cognitive 
ability [@suddendorf_Evolution_2007]. Thus, standard logic of 
Lloyd Morgan’s cannon demands that operant conditioning as the simpler 
explanation is to be accepted by default. Here, we explore the 
possibility that fitting a model in which the action of these two 
mechanisms, operant conditioning and chaining, depends 
on model parameters may reveal which process cleaners actually use. 
This is because one process may be better at predicting the observed 
variation in cleaner fish performance as a function of ecological conditions.

@@

```{r post, fig.cap="Posterior distributions for parameter values $\\gamma$, $\\eta$ and the scaling constant. We show the kernel density estimates, below the mode (black point) and the 65\\% (light blue) and 95\\% (grey)  highest posterior density interval for the three paramteres. On the left (panels a, c and c) posterior distributions for a model with   all three parameters; while on the right (b and d) a model without negative reward. Panel f shows a measure of fit for both models, namely the distribution of pseudo-$R^2$ obtained from sampling the posterior distribution of parameter values.",out.width='100%',fig.height=7}
library(data.table)
library(coda)
library(ggplot2)
library(cowplot)
library(ggdist)
source(here("../R_files/posPlots.R"))
source(here("loadData.R"))
library("jsonlite")

@@

Figure \ref{fig:post} panels a,c and e show the posterior distributions
of the parameter values of the full model, which includes both future
and negative reward. Posterior distributions show the probability
density (y axis) of the estimated parameter values (x axis). Notably,
the bulk of the posterior distribution for the parameter tuning negative
reward ($\eta$ panel e) is around zero. Thus, the shape of the posterior
and its credible intervals suggest the absence of negative reward to be
well supported by the data. As for $\gamma$, the confidence intervals
also includes zero, but the mode of the posterior is around 0.5. Thus,
we run the analysis setting $\eta$ to zero. Figure \ref{fig:post} panels
b and d, show the posterior distributions of the two parameters
estimated. Not surprisingly, under a model without negative reward, the
distribution of $\gamma$ shifts to higher values. Under the new model,
the estimates of $\gamma$ do not include zero, the distribution is
centered around $0.5$. To asses the fit of the two models we calculated
the $pseudo-R^2$ [@mcfadden_Conditional_1974], which is a standard
measure of fit for logistic regression. In that context, $pseudo-R^2$
uses the log-likelihood of the data given the model, relative to the
log-likelihood of the data given a model without covariates, as a
measure of fit. Given that our model is not a logistic regression, we
measure the $pseudo-R^2$ in relation to the log-likelihood of a model
with parameters $\gamma$ and $\eta$ set to zero. This in practice
amounts to a model that has neutral preference between the two options.
In panel f of figure \ref{fig:post} we show the distribution of
$pseudo-R^2$ calculated using samples from the posterior distributions
shown before, as well as from a model with negative reward but without
future reward. Note, $pseudo-R^2$ can have negative values, that is when
the log-likelihood of the model is lower than that of a model that
triggers neutral preferences. Even though, the peak of the three
$pseudo-R^2$ distributions is not very different, the model with only
future reward produces a distribution of $pseudo-R^2$ where more values
are positive (to the right of black line in Fig. \ref{fig:post} f). This
shows that, accounting for variation in the parameter estimates, the
model with only future reward gives a better fit to the data, despite
having one parameter less.

```{r pred, fig.cap="Observed and predicted probability of choosing a visitor. Left-hand side panel: color contour shows the prediction of the learning model using the mode of the posterior distributions of parameters recovered by the statistical analysis. Dots show the frequency of visitor choices for the 12 locations, as well as the corresponding relative cleaner abundance (x axis) and frequency of visitors leaving the station (y axis). Right-hand side panels: scatter plot of the observed and predicted probabilities of choosing a visitor for 12 locations. Line correspond to a perfect match between observed and predicted probabilities. Pseudo $R^2$ shown in the lower right corner corresponds to the fit of a model parameterized with the mode of the posterior distributions. Upper panels show prediction from a model including future and negative reward, lower panels froma  model with only future reward.",out.width='100%',fig.height=6}

source(here("manFigures.R"))
ggarrange(cont.obs.pred.both,scatter.obs.pred.both,
cont.obs.pred.gam,scatter.obs.pred.gam,
          labels=c('a','b','c','d'),common.legend=TRUE,legend = "top")
```

Figure \ref{fig:pred} shows predictions from the two models. On the
left, the colour coded contour shows the prediction of the model for
combinations of cleaner abundance (x axis) and visitor leaving
probability (y axis). Predictions were generated by running the model
using the mode of the posterior distributions as the parameter values,
and varying systematically cleaner abundance and visitor leaving
probability. Colour of the points on top of the contour show the average
frequency with which cleaner fish chose the visitor plate in their last
20 experimental trials. Each point corresponds to a location with a
combination of cleaner abundance and visitor leaving probability. Panels
on the right show how close the observed values for the probability of
choosing the visitor (y axis) are from the predictions of the models (x
axis). The line represents the perfect fit between observed and
predicted values. Predicted values were taken using the mode just as in
the left panels. On the lower right corner of the panel we show the
$pseudo-R^2$ for predictions using the mode of the posterior
distributions. Despite having one more parameter varying, the full model
gives a lower fit to the data than the model without negative reward.
The difference in $pseudo-R^2$ for the point estimates is small, but
using the whole posterior distribution shows the improvement in fit
achieved by eliminating negative reward from the model (Fig.
\ref{fig:post}f). In the supplementary material, we also show that a
model with only negative reward, excluding future rewards ($\gamma=0$,
Fig. \ref{fig:nogamma}), provides a lower fit than either of the ones
presented in the main text.

@@

We have shown that using future rewards in the learning process is the
cognitive strategy that best explains the intra-specific variation in
cleaner performance in the market task. We compared future reward with
the use of visitors behaviour as a primary source of negative reward,
and a combination of both. All three models capture well the positive
relation between visitor leaving behaviour and cleaner performance in
the market task [@triki_Biological_2019]. For both mechanisms it is
indeed necessary to have fleeing behaviour on the side of the visitors
for cleaners to develop a preference for them. The key difference
between future and negative reward emerges in the predicted relation
between cleaner relative abundance and cleaner performance. When cleaner
relative abundance is low a model with only future reward predicts low
performance in the market task regardless of the visitor leaving
probability. In contrast, a model with negative reward predicts the
highest performance with low relative cleaner abundance, particularly
together with high probability of visitor leaving (Fig. \ref{fig:pred}).
Low cleaner abundances mean the market has an excess of demand for
cleaning services. In the model, this translates to a cleaning station
that is frequently full. Thus, when visitors leave, it's likely that the
cleaner will have access to another client, and there will not be much
difference in future reward between choosing a visitor and a resident.
Thus, estimation of future reward will not develop a preference for the
visitor in this conditions, because it matches the state of the market.
On the other hand, the effect of negative reward is unaffected by this
market situation. In a full station, cleaners will get more often the
heterotypic (resident-visitor) option and will develop a preference for
the visitor faster. Note that when cleaner abundance is high situations
with a resident and a visitor will be so rare that neither mechanism
will be very efficient in generating a preference for visitors. When
such choice arises it is still be best to choose the visitor; however,
the learning machinery will not be able to develop this preference
efficiently. Overall, estimation of future reward is then the cognitive
mechanism that allows cleaners to adaptively respond conditions to the
ecological conditions of the biological market.

The cognitive mechanism that our analysis highlights does not require a
direct flow of information between the cleaner and its social
environment. The adaptive responses to local conditions in cleaners is
not only evidenced by their performance in experiments, but also by
changes in their brain anatomy. Indeed, individual cleaners that
performed in the laboratory test according to what is best at the
location in which they were caught had larger forebrains harbouring more
cells than individuals whose performance mismatched locally optimal
decision-rules [@triki_Brain_2020]. This suggests that despite the
overall behavioural adjustment, evidenced here and elsewhere
[@triki_Biological_2019; @triki_Decrease_2018], there are some
cleaners better than others at adjusting their behaviour. Triki *et al.*
[@triki_Brain_2020] refer to these as socially competent cleaners. Social
competence refers to the ability to optimise social behaviour depending
on the available social information [@taborsky_Social_2012a;
@bshary_Cooperation_2015]. However, we have shown here that cleaners
are able to adjust their preference according to the social environment
without using any direct information about the social environment.
Instead cleaners seem to adjust their preference using information from
the visitation patterns of their clients. These patters are of course
determined by the feedback with the cleaner's social environment
(relative abundance). Our models, however, do not include the
possibility of cleaners responding to their relative abundance directly.
The mechanism we present is a general way to learn from the environment
in both asocial and social situations. It would be interesting to
contrast this learning mechanism with one tailored to a social
situation, and see whether such mechanism fits the data better. More
generally, an important challenge ahead is to understand which cognitive
mechanisms evolved exclusively for social life.

One potential difference between cleaners that adjusted adaptively, like
the ones reported in Triki *et al.* [@triki_Brain_2020], is configurational learning or
chunking. The original analysis of the learning model showed that agents
needed two mechanism to solve the market task; to account for the
future, and to assign value to the combination of clients present in the
station [@quinones_Reinforcement_2019]. The latter is referred to as
configurational learning or chunking [@sutherland_Configural_1989;
@miller_Magical_1956]. Theoretical models suggest that the lack of
configurational learning is related to a low frequency with which
stimuli are encountered together [@kolodny_Evolution_2014;
@prat_Role_2021]. Thus, cleaner relative abundance could be related to
this mechanisms as well. Given that the lack of either of these two
mechanisms yields neutral preferences [@quinones_Reinforcement_2019],
we can't, based on the current analysis, distinguish between the two
possibilities. One potential way to distinguish between the two would be
to use an experimental set up designed to test configuration learning on
cleaners with high and low performance in the market task
[@woodbury_Learning_1943; @sutherland_Configural_1989].

>>>>

---
title: Combination of field and experimental data together with computational models
  reveal cognitive mechanisms behind cleaner fish behaviour
author:
- Andrés E. Quiñones^[Institute of Biology, University of Neuchâtel, Neuchâtel, Switzerland
  - andresqp@gmail.com]
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  word_document: default
  pdf_document: default
fig_caption: yes
fontsize: 12pt
header-includes:
- \usepackage{float}
- \usepackage{setspace}\doublespacing
- \linespread{2}
- \usepackage{lineno}
- \linenumbers
- \newcommand{\beginsupplement}{ \setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}}
  \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
keep_tex: yes
linespread: 2
bibliography: Cleanerlearning.bib
style: M:\Zotero\styles\behavioral-ecology-and-sociobiology.csl
spacing: double
abstract: |
  While it is generally straightforward to quantify individual performance in cognitive experiments, identifying the underlying cognitive processes remains a major challenge. Often, different mechanistic underpinnings may yield similar performances, and Lloyd Morgan’s cannon warrants acceptance of the simpler explanation. Alternatively, when the different mechanisms interact with environmental conditions, variation in performance across environments might allow to statistically infer the mechanism responsible. However, to do this it is necessary to have quantitative predictions for the different candidate mechanisms. Here, we use a set of computational models to get quantitative predictions from alternative learning mechanisms, and using Bayesian statistics fit the model parameters to performance data. We used experimental data on performance in an ephemeral reward task by wild-caught cleaner fish *Labroides dimidiatus*, as well as cleaner and client fish densities from the locations of capture. The task can in principle be solved by estimating future consequences of an action, or by perceiving the removal of the ephemeral reward as psychological punishment (negative reinforcement). We found that a model where cleaners only estimate the future consequences of their actions explains best which cleaner relative abundances cause the fish to develop a preference for an ephemeral food source. This model also yields performances that can be considered the result of locally optimal decision-rules, in contrast to the negative reinforcement model.  We argue that the combination of computational models with data is a powerful tool to infer the mechanistic underpinning behind cognitive performance. 
keywords: learning, behaviour, cleaners, bayesian statisitics
---

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning = FALSE,message=FALSE,cache=TRUE)
library(here)
```

Cooperation among individuals of the same or different species is
ubiquitous in nature. Whenever such interactions involve some degree 
of partner choice, biological market theory predicts that the exchange 
rate of goods and/or services traded adjusts to the law of supply and demand 
[@noe_Biological_1995a]. While biological market models have largely focused 
on evolutionary time scales [@noe_Biological_1994a;@johnstone_Mutualism_2008a] 
$\color{red}{\text{Hoekstra xxxx}}$, supply and demand conditions, which typically 
depend on the abundance of the species involved, certainly vary in time 
and space. Therefore, natural selection should 
favour the ability to flexibly adjust decisions and behavioural output to 
current market conditions. Indeed, such adjustments have been repeatedly 
documented [@axen_Signalling_1996]. In animals, an obvious general candidate 
mechanism for the strategic adjustment is the cognitive machinery. However, 
it is not clear which cognitive mechanisms allow individuals to adjust their 
behaviour to the varying market conditions. As adjustments to changes in 
market conditions have been documented in a great variety of species with 
highly variable brain anatomies, the question arises to what extent mechanisms 
beyond basic associative learning may be involved.

One example of strategic adjustment in a biological market is the marine
cleaning mutualism involving the cleaner fish *Labroides dimidiatus* and
'client' fish. Clients offer themselves as food patches for cleaners to
remove ectoparasites, which gives the cleaners calories and the clients
improved health [@waldie_LongTerm_2011;
@ros_Does_2011;@triki_Effects_2016]. As cleaners have about 2000
interactions per day [@grutter_Relationship_1995], they regularly face
the situation that two clients seek cleaning simultaneously. In such
conditions, it is important to distinguish between 'resident' and
'visitor' client species [@bshary_Cleaner_2001a]. The residents have
small home ranges and hence access to only the local cleaner, while the
visitors have large home ranges and hence access to two or more
cleaners. Thus, visitors have the option to switch to another cleaner if
being made to wait, while residents must wait for inspection. Indeed,
visitors have been observed to use their partner choice options in that
way [@bshary_Choosy_2002], which may explain why cleaners give
visitors service priority in a field study in the Red Sea
[@bshary_Cleaner_2001a]. More generally, cleaners apparently prefer to
interact with visitors [@triki_Marine_2021]. Such a preference may be
linked to a client's value as a food patch; visitors are on average
larger than residents [ @bshary_Cleaner_2001a], and size is positively
correlated with parasite load [@grutter_Relationship_1995]. However, 
early laboratory experiments showed that leaving causes service 
priority independent of client value [@bshary_Asymmetric_2002]. 
The experiments consisted of asimultaneous two-choice task, 
where cleaners had to decide which of two
Plexiglas plates, which offered equal amounts of food, to approach first
and eat mashed prawn of it. One plate was made to behave like a
resident, i.e. it would remain until the cleaner had eaten off it. The
other plate was made to behave like a visitor, i.e. it would be removed
if not inspected first. Cleaners learned to prefer the visitor plate and
hence obtained the double amount of food [@bshary_Asymmetric_2002].
Adult cleaner fish outperform various primates as well as rats and
pigeons in this original version of the biological market task
[@salwiczek_Adult_2012; @zentall_Early_2017]; only African grey
parrots solve this task as well [@pepperberg_Can_2014]. As the correct 
solution must be learned, and primates are known to readily distinguish 
between one and two reward units if presented in more conventional ways 
$\color{red}{\text{ref}}$, what makes the task difficult for most non-cleaners?

Recent research on the cognitive tool kit needed to solve an ephemeral 
reward task has used a broad approach as proposed by Shettleworth 
[@shettleworth_Cognition_2009], 
who defined cognition as including all ways in which animals take in information 
through the senses, process, retain and decide to act on it. Indeed, perception 
of relevant cues is of major importance for individual performance; 
i.e. what relevant information is encoded in the design of the two 
plates (colour versus size; [@wismer_Cuebased_2019]), or whether 
the information is in the colour of
the food [@pretot_Comparative_2021; @pretot_Comparing_2016], and
whether the food is visible or hidden/released only when the subjects
made choices on a touch screen [@pretot_Factors_2016]. However, identifying 
cues as salient is not the only challenge of the ephemeral reward task, 
as revealed by proximate learning models. Such models allow varying the 
cognitive tool kit and to evaluate which minimal kit is necessary to solve 
the task at hand (e.g. [@dubois_Model_2021]). Applied to the ephemeral 
reward task, learning models showed that basic reinforcement learning does 
not suffice to solve the task [@prat_Modelling_2022; @quinones_Reinforcement_2019]. 
This is particularly so when models assume the more complex natural 
situation in which cleaners not only face resident-visitor pairs but also 
visitor-visitor and resident-resident pairs, as well only a resident or 
only a visitor. To be able to give visitors priority over residents, 
cleaners need to be able to assess a client’s value separately for the 
three possible forms of encounter (alone, paired with a fish with the 
same strategic option, paired with a fish with the alternative strategic option) 
[@quinones_Reinforcement_2019]. The ability to keep such separate scores 
for the same stimulus has been termed configurational learning, 
or chunking, or segmentation (see references in [@prat_Modelling_2022]).

To solve the ephemeral reward task, cleaners also need to account for 
the future consequences of current decisions. In the model by 
Quiñones *et al.* [@quinones_Reinforcement_2019], this could be achieved in 
two non-mutually exclusive ways: through low temporal discounting of 
future effects, also termed ‘chaining’ [@enquist_Power_2016]; and/or 
through perceiving a visitor client leaving as psychological punishment 
(i.e. as a negative reinforcer). Given that a negative reinforcer is part 
of the ubiquitous learning mechanism termed operant conditioning 
[@thorndike_Animal_1898; @skinner_Behavior_1938], this explanation 
could be considered the simpler one. Furthermore, accounting for the 
future effects of an action is considered an advanced cognitive 
ability [@suddendorf_Evolution_2007]. Thus, standard logic of 
Lloyd Morgan’s cannon demands that operant conditioning as the simpler 
explanation is to be accepted by default. Here, we explore the 
possibility that fitting a model in which the action of these two 
mechanisms, operant conditioning and chaining, depends 
on model parameters may reveal which process cleaners actually use. 
This is because one process may be better at predicting the observed 
variation in cleaner fish performance as a function of ecological conditions.

@@

```{r post, fig.cap="Posterior distributions for parameter values $\\gamma$, $\\eta$ and the scaling constant. We show the kernel density estimates, below the mode (black point) and the 65\\% (light blue) and 95\\% (grey)  highest posterior density interval for the three paramteres. On the left (panels a, c and c) posterior distributions for a model with all three parameters; while on the right (b and d) a model without negative reward. Panel f shows a measure of fit for both models, namely the distribution of pseudo-$R^2$ obtained from sampling the posterior distribution of parameter values.",out.width='100%',fig.height=6}
library(data.table)
library(coda)
library(ggplot2)
library(cowplot)
library(ggdist)
source(here("../R_files/posPlots.R"))
source(here("loadData.R"))
library("jsonlite")

@@

Figure \ref{fig:post} panels a,c and e show the posterior distributions
of the parameter values of the full model, which includes both future
and negative reward. Posterior distributions show the probability
density (y axis) of the estimated parameter values (x axis). Notably,
the bulk of the posterior distribution for the parameter tuning negative
reward ($\eta$ panel e) is around zero. Thus, the shape of the posterior
and its credible intervals suggest the absence of negative reward to be
well supported by the data. As for $\gamma$, the confidence intervals
also includes zero, but the mode of the posterior is around 0.5. Thus,
we run the analysis setting $\eta$ to zero. Figure \ref{fig:post} panels
b and d, show the posterior distributions of the two parameters
estimated. Not surprisingly, under a model without negative reward, the
distribution of $\gamma$ shifts to higher values. Under the new model,
the estimates of $\gamma$ do not include zero and the distribution is
centered around $0.5$. In panel f of figure \ref{fig:post} 
we show the distribution of $pseudo-R^2$ calculated using 
samples from the posterior distributions shown before, 
as well as from a model with negative reward but without
future reward. Note, $pseudo-R^2$ can have negative values, that is when
the log-likelihood of the model is lower than that of a model that
triggers neutral preferences. Even though, the peak of the three
$pseudo-R^2$ distributions is not very different, the model with only
future reward produces a distribution of $pseudo-R^2$ where more values
are positive (to the right of black line in Fig. \ref{fig:post} f). This
shows that, accounting for variation in the parameter estimates, the
model with only future reward gives a better fit to the data, despite
having one parameter less.

```{r pred, fig.cap="Observed and predicted probability of choosing a visitor. Left-hand side panel: color contour shows the prediction of the learning model using the mode of the posterior distributions of parameters recovered by the statistical analysis. Dots show the frequency of visitor choices for the 12 locations, as well as the corresponding relative cleaner abundance (x axis) and frequency of visitors leaving the station (y axis). Right-hand side panels: Variation of the predicted probabilites  of choosing a visitor and their observed values for 12 locations. Circles show the mean prediction for each location from 100 samples taking from the posterior distribution. Thick and thin bars show the 66 and 95\\% credible interval taken from those posterior samples. Squares show the predictions used for panel on the righ-hand side. Colour denotes different locations (see Empirical data section). Black line correspond to a perfect match between observed and predicted probabilities. Upper panels show prediction from a model including future and negative reward, lower panels from a  model with only future reward.",out.width='100%',fig.height=5}

source(here("manFigures.R"))
ggarrange(cont.obs.pred.both,scatter.obs.pred.both,
cont.obs.pred.gam,scatter.obs.pred.gam,
          labels=c('a','b','c','d'),common.legend = F,
  heights=c(1.25,1))

Figure \ref{fig:pred} shows predictions from the two models. On the
left, the colour coded contour shows the prediction of the model for
combinations of cleaner abundance (x axis) and visitor leaving
probability (y axis). Predictions were generated by running the model
using the mode of the posterior distributions as the parameter values,
and varying systematically cleaner abundance and visitor leaving
probability. Colour of the points on top of the contour show the average
frequency with which cleaner fish chose the visitor plate in their last
20 experimental trials. Each point corresponds to a location with a
combination of cleaner abundance and visitor leaving probability. Panels
on the right show how close the observed values for the probability of
choosing the visitor (y axis) are from the predictions of the models (x
axis). The black line represents the perfect fit between observed and
predicted values. Bars show the variation in predicted values using 100 
samples from the posterior distribution. Despite having one more 
parameter varying, the full model
gives a lower fit to the data than the model without negative reward.
In the supplementary material, we also show predictions from 
model with only negative reward, excluding future rewards ($\gamma=0$,
Fig. \ref{fig:nogamma}), provides a lower fit than either of the ones
presented in the main text.

@@

We asked which if two potential cognitive mechanisms 
– chaining of events, negative reinforcement, or a combination of both – 
is the more likely mechanism to explain intra-specific variation in cleaner 
performance in the ephemeral reward task as a function of ecological 
conditions in the wild. To evaluate the merits of the alternative mechanisms, 
we considered performance in the task to depend on decision-rules that 
cleaners had developed in their natural environment: individuals that 
solved the task did so because they already had a general preference for 
visitors, and applied this rule once being familiar with the task.  
While all three models capture well the positive relation between visitor 
leaving behaviour and cleaner performance in
the market task [@triki_Biological_2019], only the chaining mechanism 
predicts that cleaner performance in the task should be low if their 
local abundance in the wild relative to clients is low. Indeed, such low 
performance is predicted regardless of the visitor leaving
probability. In contrast, a model with negative reward predicts the
highest performance in the ephemeral reward task when relative 
cleaner abundance is low, particularly
together with high probability of visitor leaving (Fig. \ref{fig:pred}).
Low cleaner abundances mean the market has an excess of demand for
cleaning services. In the model, this translates to a cleaning station
that is frequently full. Thus, when visitors leave, it's likely that the
cleaner will have access to another client in the next step. Therefore, 
there will not be much difference in future reward between choosing 
a visitor and a resident, and cleaners will not develop a preference for the
visitor in these conditions. On the other hand, the effect of negative reward 
would be opposite as in a full station, cleaners will get more often the
heterotypic (resident-visitor) option and will develop a preference for
the visitor faster. Note that when cleaner abundance is high situations
with a resident and a visitor will be so rare that neither mechanism
will be very efficient in generating a preference for visitors. When
such choice arises it is still be best to choose the visitor; however,
the learning machinery will not be able to develop this preference
efficiently. Overall, estimation of future reward is then the cognitive
mechanism that allows cleaners to adaptively respond to the
ecological conditions of the biological market.

Our results support chaining as the more plausible mechanism behind the 
inter-specific variation in cleaners' performance, which is also 
the mechanism that allows them to respond more adaptively. This could
suggest chaining to be under positive selection. In the standard ephemeral
reward task, however, food amount and quality are equal between the 
two plates. An given that models resemble this, choosing visitors over 
residents in low cleaner abundances does not diminish the 
amount of food/reward in the the models. 
On the other hand, a cleaner committed to visitors might find more difficult 
to include in the decision making other variables, like parasite load 
and/or mucus quality, that do predict the amount of food in the field 
[@roche_Client_2021]. In nature and in laboratory experiments, cleaners prefer 
higher quality mucus and more food in the absence of market effects 
(i.e. in simultaneous 2-choice options, 
[@triki_Fluctuations_2019;@grutter_Cleaner_2004]), and they use size as a 
proxy for food availability [@wismer_Cuebased_2019;@grutter_Does_2005]. 
How important this limitation is will determine whether chaining yields a higher
fitness for cleaners.

Individual cleaners that performed in the laboratory test according to what 
is best at their site of capture (low performance when living at low densities, 
high performance when living at high densities) had larger forebrains 
harbouring more cells than individuals whose performance mismatched locally 
optimal decision-rules [@triki_Brain_2020]. Triki *et al.*  refer to the 
former as socially competent cleaners. Social competence refers to the 
ability to optimise social behaviour depending on the available social 
information [@taborsky_Social_2012a;
@bshary_Cooperation_2015;@varela_Correlated_2020]. Our analyses yielded no 
evidence that the difference in social competence with respect to the local 
market conditions and associated brain morphology are due to the mechanism 
used to incorporate future consequences. It could have been conceivable that 
high performing individuals from low-density sites might use negative 
reinforcement instead of chaining, but in that case negative reinforcement 
should have explained part of the data. Configurational learning or 
chunking [@sutherland_Configural_1989;@miller_Magical_1956], the second component 
necessary to solve the ephemeral reward task [@quinones_Reinforcement_2019], was 
not varied in the models analized here. However, while chunking tendencies 
should vary to allow individuals to adapt to local conditions 
[@prat_Modelling_2022;@kolodny_Evolution_2014], systematic differences 
in individual chunking tendencies would not explain how socially 
competent decisions vary as a function of relative abundance. 
Therefore, it remains currently unclear what cell-demanding 
mechanisms may cause variation in social competence that 
translates into site-specific variation in performance in the 
ephemeral reward task.

The model is inspired by the general processes of associative learning
where short term rewards are translated to decision making; thus, it 
ignores alternative channels of information that could be relevant in 
market-like situations. For example, the model does not investigate whether 
cleaners actually assess the frequency of client visits or a mean frequency 
of visitors leaving. The updating learning mechanism for the development 
of preferences works on a trial by trial basis. In the model, cleaners 
do not need to assess the actual state of the market, *i.e.* their density, 
the density of residents and visitors, and client visitation rate as an 
indicator of demand. They only need to assess short-term consequences of 
own decisions on food intake and chain them. Also, for the sake of 
simplicity, the model ignores the process by which cleaners discriminate 
residents and visitor clients. A model that accounts for this discrimination
probably involves the development of preferences for morphological or 
behavioural features that are statistically associated with visitors 
or residents. For example, visitors are on average larger larger 
than residents [@bshary_Cleaner_2001a], and are less likely to chase 
cleaners in response to the latter cheating by taking a bite of 
mucus [@bshary_Asymmetric_2002]. Given these associations, chaining might 
produce the decision-rule “chose the larger client and/or less aggressive 
client”, which does not help in the standard ephemeral reward task.
